---
title: "Preliminary Analyses"
author: "Jacob Bukoski"
date: "8/25/2020"
output: html_document
---

```{r, echo = FALSE, message = FALSE}
library(tidyverse)
library(ggthemes)
```


## Update on FIA work - August 25, 2020

</br>

  1. Gathered data for single burn and reburn plots from the periodic inventories, the R5 inventories and the annual inventories (with a ton of Jeremy's help - thank you!)
  2. Began linking the data and trying to calculate biomass to begin looking at effects and differences in data
  3. Lost a lot of plots due to the absence of either a pre- or post-fire measurement, mostly the pre-fire measurements
  4. For reburn plots
      * Most pre-fire measurements come from periodic 1990s database
  5. For single burn plots
      - Most pre-fire measurements come from R5 database. Very few pre-burn visits were found in the annual data - most are post-fire.
  6. Have calculated biomass for the plots that are available, took some time to understand different designs and identify different variables
      - Calculating biomass estimate from each plot in Mg C ha^-1
      - Filtered all tree records to just those that were > 1 in diameter (to align periodic and R5 with annual)
  7. For plots with pre-burn and post-burn measurements of biomass, we had:
      - 154 reburn plots (78 softwood, 87 hardwood)
      - 161 single burn plots (98 softwood, 78 hardwood)
  8. Calculated:
      - pre-fire biomass (Mg C ha-1)
      - post-fire biomass (Mg C ha-1)
      - percent change (100 * (post - pre / pre))
      - Post-Fire to Burn time (yrs)
      - Burn to Pre-Fire time (yrs)
  9. Things get a bit wonky with the percent change, so I was looking at things in terms of loss (i.e., precent change > -100 & < 0) and gain (percent change > 0)

</br>

```{r, echo = FALSE, message = FALSE} 

raw_dat <- read_csv("~/Dropbox/manuscripts/caliFIA/analysis/data/processed/conf_plots_biomass.csv")

# Calculate how many reburn and single plots we have pre and post-fire data for.

counts <- raw_dat %>%
  group_by(PLOT_FIADB) %>%
  filter("1_Softwoods" %in% SWHW) %>%
  ungroup() %>%
  group_by(BURN) %>%
  summarize( n = n_distinct(PLOT_FIADB) )

```

### Difference in mean loss by burn type

</br>

First for plots that exhibited losses in biomass. Perhaps a significant difference...

```{r, echo = FALSE, warning = FALSE, message = FALSE}

raw_dat %>%
  filter(perc_diff < 0) %>%
  select(PLOT_FIADB, BURN, perc_diff) %>%
  distinct() %>%
  group_by(BURN) %>%
  summarize(diff_avg = mean(perc_diff),
            diff_se = plotrix::std.error(perc_diff))

```

</br>

Now for plots that exhibited gains in biomass.

```{r, echo = FALSE, warning = FALSE, message = FALSE}

raw_dat %>%
  filter(perc_diff > 0) %>%
  select(PLOT_FIADB, BURN, perc_diff) %>%
  distinct() %>%
  group_by(BURN) %>%
  summarize(diff_avg = mean(perc_diff),
            diff_se = plotrix::std.error(perc_diff))

```

</br>

### Plotting histograms of the percent differences by burn class

</br>

Doesn't seem to be much visual difference?

```{r, echo = F, warning = F, message = F}

raw_dat %>%
  group_by("PLOT_FIADB") %>%
  #filter("1_Softwoods" %in% SWHW) %>%
  ungroup() %>%
  select(BURN, perc_diff) %>%
  distinct() %>%
  ggplot() +
  facet_grid(. ~ BURN) +
  geom_histogram(aes(perc_diff), bins = 16) +
  xlim(c(-100, 10)) +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
```

### Plotting pre (red) & post (blue) biomass ~ burn class

</br>

Red is pre-fire visit whereas blue is post-fire visit biomass. No filter on SWHW or biomass gain/loss.

Doesn't seem to be much difference there...

```{r, echo = FALSE}

raw_dat %>%
  select(PLOT_FIADB, BURN, PRE, POST, perc_diff) %>%
  distinct() %>%
  #filter(perc_diff < 0) %>%
  ggplot() +
  geom_boxplot(aes(x = BURN, y = PRE), col = "red", alpha = 0.1, width = 0.3) +
  geom_boxplot(aes(x = BURN, y = POST), col = "blue", alpha = 0.1, width = 0.3) +
  theme_tufte() +
  ylab("Biomass C (Mg/ha)") +
  xlab("Burn Class") +
  theme(panel.grid.major = element_blank())

```

### Repeating the same plot, but for softwoods and only those that experienced losses

</br>

Perhaps a trend of slightly greater loss in biomass post-fire (blue plots) for single-burn plots.

```{r, echo = FALSE}

raw_dat %>%
  distinct() %>%
  filter(perc_diff < 0) %>%
  group_by(PLOT_FIADB) %>%
  filter("1_Softwoods" %in% SWHW) %>%
  select(PLOT_FIADB, BURN, PRE, POST, perc_diff, OWNGRPCD) %>%
  distinct() %>%
  ungroup() %>%
  ggplot() +
  geom_boxplot(aes(x = BURN, y = PRE), col = "red", alpha = 0.1, width = 0.3) +
  geom_boxplot(aes(x = BURN, y = POST), col = "blue", alpha = 0.1, width = 0.3) +
  theme_tufte() +
  ylab("Biomass C (Mg/ha)") +
  xlab("Burn Class") +
  theme(panel.grid.major = element_blank())

```

### Looking at the differences in pre- vs. post-fire biomass by ownership group code & for softwoods only.

</br>

Post-fire biomass is definitely lower on state and private lands... bit of a wash on federal lands.

```{r, echo = FALSE}

raw_dat %>%
  distinct() %>%
  filter(perc_diff < 0) %>%
  group_by(PLOT_FIADB) %>%
  filter("1_Softwoods" %in% SWHW) %>%
  select(PLOT_FIADB, BURN, PRE, POST, perc_diff, OWNGRPCD) %>%
  distinct() %>%
  ungroup() %>%
  ggplot() +
  facet_grid(. ~ OWNGRPCD) + 
  geom_boxplot(aes(x = BURN, y = PRE), col = "red", alpha = 0.1, width = 0.3) +
  geom_boxplot(aes(x = BURN, y = POST), col = "blue", alpha = 0.1, width = 0.3) +
  theme_bw() +
  ylab("Biomass C (Mg/ha)") +
  xlab("Burn Class") +
  theme(panel.grid.major = element_blank())

```


### Running a few perhaps-too-early models

</br>

Data structure is as follows:

  - POST = post-fire biomass
  - PRE = pre-fire biomass
  - BURN = burn class
  - perc_diff = percent difference between POST and PRE
  - TimeSinceFire = number of yrs between post-visit and fire
  - TimeBeforeFire = number of yrs between pre-visit and fire

</br>

Example head of data:

```{r, echo = FALSE}

mdl_dat <- raw_dat %>%
  filter(perc_diff <= 0) %>%
  select(PLOT_FIADB, POST, PRE, BURN, perc_diff, TimeSinceFire, TimeBeforeFire, OWNGRPCD) %>%
  distinct()

head(mdl_dat)

```

</br>

### Run a simple model on the data, BOTH hardwood and softwood

</br>

Total of 230 distinct plots.

Don't find any effect for burn class. Find strong effects for pre-fire biomass, ownership group, SWHW class.

Weak effect for single burn vs. reburn

```{r, echo = FALSE}

mdl1 <- glm(POST ~ BURN + PRE + TimeSinceFire + TimeBeforeFire + OWNGRPCD, data = mdl_dat)

summary(mdl1)
            
```

</br>

### Repeat for only softwoods

Total of 123 plots included in the model.

Seems to find a weak effect for burn class. 
Not particularly convincing and things are likely to change with additional data or changing calculations.

```{r, echo = FALSE}

mdl_dat <- raw_dat %>%
  group_by(PLOT_FIADB) %>%
  filter("1_Softwoods" %in% SWHW) %>%
  ungroup() %>%
  filter(perc_diff <= 0) %>%
  select(PLOT_FIADB, POST, PRE, BURN, perc_diff, TimeSinceFire, TimeBeforeFire, OWNGRPCD) %>%
  distinct()

mdl1 <- glm(POST ~ BURN + PRE + TimeSinceFire + TimeBeforeFire + OWNGRPCD, data = mdl_dat)

summary(mdl1)
            
```

</br>

### Questions/points for further discussion

  - Would appear to be pretty messy with variable results depending on what's included... but something might be there.
  - Clear that there appears to be a strong effect for both pre-fire biomass and ownership group, will want to keep these in the modeling.
  - Is there a better model form to be using?
  - How do we refine the comparisons such that there would be stronger effects?
      - Likely need to go down to the condition level - ownership is a little bit messy in the table (multiple entries for some plots, though certainty the minority)
      - Can look by region?
      - Better calculations of biomass?
  - Stronger effects for other variables?
      - Seedlings?
      - Number of species on plots?
      - Dead wood?
      
</br>
