---
title: "county_density_plots"
author: "Evan Patrick"
date: "7/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Input a set of R5 plots from the FIADB and generate a heat map of plot density based on county. 

```{r}
# trouble installing XML package (needed for tmap) - run following code
#install.packages("XML", type="binary")
library(rFIA)
library(RSQLite)
library(tidyverse)
library(sp)
library(rgdal)
library(rgeos)
library(maptools)
library(tmap)

# Set working dir
setwd("~/Summer 2020/FIA/rFIA")

```

Read in county shapefile

```{r}
CAcounty = readOGR("./data/ca-county-boundaries/CA_Counties/CA_Counties_TIGER2016.shp")

```

Get CountyFP from plots. First read in plot data, FIADB data, then join to get CountyFP

```{r}
# read in reburn plot data and filter out repeated plot entries
confirmed_reburn_plots <- read.csv("./data/confirmed_plots.csv") %>% 
  rename(PLOT = PLOT_FIADB) %>% distinct(PLOT, .keep_all = TRUE)

# Get filename of db and read into SQLite 
filename <- "./data/PNW_PUBLIC_SQLite.db"
sqlite.driver <- dbDriver("SQLite")
db <- dbConnect(sqlite.driver, dbname = filename)
# pull out plot table from larger db
pnwFIADBplot <- dbReadTable(db, "PLOT")
# read condition plot table 
pnwFIADBcond <- dbReadTable(db, "COND")

# join reburn plot data with FIADB plot and condition tables
plot_join <- confirmed_reburn_plots %>% left_join(pnwFIADBcond, by = "PLOT") %>%
  distinct(PLOT, .keep_all = TRUE) 
# filter out data that produced NAs, filter out duplicate plot entries
plot_join_filt <- filter(plot_join, !is.na(PLOT), !is.na(COUNTYCD))

# figure out which plots are missed 
missed_plots <- anti_join(confirmed_reburn_plots, plot_join_filt, by = "PLOT") %>%
  distinct(PLOT, .keep_all = TRUE)
```

Probably the best way to go about this is to add a column to the CAcounty@data which includes the count of plots within each county. 

```{r}
# create a list of county FP 
CountyData <- CAcounty@data %>% as.data.frame()
# CountyFP and CountyCD are different - need to add a column 
# that converts to county name. Created COUNTYFP_CA.csv for this. 
COUNTYCD_Name <- read.csv('./data/COUNTYCD_CA.csv')
# join with plot_join_filt - we lose 6 plots here b/c COUNTYCD
# does not line up with FIADB user guide
plots_per_county <- plot_join_filt %>% add_count(COUNTYCD) %>% 
  left_join(COUNTYCD_Name, by = "COUNTYCD") %>% filter(!is.na(NAME))
CountyJoin <- CountyData %>% left_join(plots_per_county, by = 'NAME') %>%
  distinct(NAME, .keep_all = TRUE)
CountyJoin$n <- replace_na(CountyJoin$n, 0)
# add 'n' column back to CAcounty data 
CAcounty@data$n <- CountyJoin$n
```

Now plot the CA counties using n as color density

```{r}

# using tmap function qtm, you can display plot density easily
qtm(CAcounty, "n", fill.n = 8)
```

Now to split up the data based on forest type and ownership

```{r}
# get forest type reference table from FIADB
ForTypRef <- dbReadTable(db, "REF_FOREST_TYPE") %>% rename(FORTYPCD = REF_FORTYPCD)
# recreated Ownership type reference table from FIADB User Guide
OWNGRPCDref <- matrix(c(11, "National Forest.", 12, "National Grassland and/or Prairie.", 13, "Other Forest Service Land.", 21, "National Park Service.", 22, "Bureau of Land Management.", 23, "Fish and Wildlife Service.", 24, "Departments of Defense/Energy.", 25, "Other Federal.", 31, "State including State public universities.", 32, "Local (County, Municipality, etc.) including water authorities.", 33, "Other non-federal public.", 46, "Undifferentiated private and Native American."), ncol = 2, byrow = TRUE) %>% as.data.frame() %>% rename( "OWNCD" = "V1", "Description" = "V2") 
OWNGRPCDref$OWNCD <- as.numeric(as.character(OWNGRPCDref$OWNCD))

# join reference tables with plots_per_county
plots_per_county_ref <- left_join(plots_per_county, ForTypRef, by = "FORTYPCD") %>%
  left_join(OWNGRPCDref, by = "OWNCD") %>% filter(!is.na(SWHW))
# this drops another 12 entries 


# split into hardwood and softwood
plots_per_county_hardwood <- filter(plots_per_county_ref, SWHW == "2_Hardwoods")
plots_per_county_softwood <- filter(plots_per_county_ref, SWHW == "1_Softwoods")

# get count of plots per county and add county name 
ppc_hardwood_count <- plots_per_county_hardwood %>% add_count(COUNTYCD) 
ppc_softwood_count <- plots_per_county_softwood %>% add_count(COUNTYCD) 
  
# join hardwood data with county data and add to CAcounty shapefile
hardwoodCountyJoin <- CountyData %>% left_join(ppc_hardwood_count, by = 'NAME') %>%
  distinct(NAME, .keep_all = TRUE)
hardwoodCountyJoin$n <- replace_na(hardwoodCountyJoin$n, 0)
CAcounty@data$hardwood_n <- hardwoodCountyJoin$n

# join softwood data with county data and add to CAcounty shapefile
softwoodCountyJoin <- CountyData %>% left_join(ppc_softwood_count, by = 'NAME') %>%
  distinct(NAME, .keep_all = TRUE)
softwoodCountyJoin$n <- replace_na(softwoodCountyJoin$n, 0)
CAcounty@data$softwood_n <- softwoodCountyJoin$n

# plot hardwood and softwood counts per county
qtm(CAcounty, 'hardwood_n')
qtm(CAcounty, 'softwood_n', fill.n = 9)

```

