---
title: "Historical Fire"
author: "Jacob Bukoski"
date: "6/17/2020"
output: html_document
---

### Examining database that Jeremy sent

Historical fire data

```{r}

library(readxl)
library(RSQLite)
library(tidyverse)

filename <- "../data/TeamPotts1.db"
sqlite.driver <- dbDriver("SQLite")
db <- dbConnect(sqlite.driver, dbname = filename)

dbListTables(db)

db_COND_LB <- dbReadTable(db, "COND_LVL_LOOK_BACKS_PNW")
db_PLOT_LB <- dbReadTable(db, "PlotLevelFireDisturbPNWFromAnnualVisitLookbacks")

```

```{r}

# Jeremy's query

q1_jsf <- db_PLOT_LB %>%
  filter(ForestAtLastPostDistVisit == -1) %>%
  select(STATECD, PLOT_FIADB, DSTRBYR1) %>%
  group_by(STATECD, PLOT_FIADB) %>%
  mutate(n = n()) %>%    # This is the problematic line: should be n_distinct(DSTRBYR1) rather than n()
  #mutate(n = n_distinct(DSTRBYR1)) %>%
  filter(n >= 2)         # Without using n_distinct above, it counts multiple conditions and multiple fire types (e.g., n = 2 on plots with only one fire, but both dstrbcd 31 & 32 were recorded)

q2_jsf <- q1_jsf %>%
  select(STATECD, PLOT_FIADB, n) %>%
  group_by(STATECD) %>%
  summarise(n2 = n_distinct(PLOT_FIADB))

```

```{r}

all_dat <- db_PLOT_LB %>%
  left_join(db_COND_LB)

#---

q1 <- all_dat %>%
  filter(STATECD == 6, ForestAtLastPostDistVisit == -1) %>%
  select(STATECD, PLOT_FIADB, INVYR, DSTRBCD1, CONDID, DSTRBYR1, Source, ForestAtLastPostDistVisit) %>%
  group_by(PLOT_FIADB) %>%    # Group by additional criteria to better identify unique plots, some plots have 
  #mutate(n = n()) %>%
  mutate(n = n_distinct(DSTRBYR1)) %>%
  filter(n >= 2) %>%
  arrange(PLOT_FIADB, DSTRBYR1, INVYR)

length(unique(q1$PLOT_FIADB))

q2 <- q1 %>%
  group_by(PLOT_FIADB) %>%
  mutate(n2 = n_distinct(interaction(PLOT_FIADB, DSTRBYR1))) %>%
  filter(n2 >= 2) %>%
  select(-ForestAtLastPostDistVisit, -n) %>%
  mutate(TimeBtwnFires = max(DSTRBYR1) - min(DSTRBYR1))   # Calculate time between fires

n_distinct(q2$PLOT_FIADB)

# Need to handle 9999 plots
# Need to resolve plots with short times between fires

```

### Read in excel table with FIA plot and fire perimeter intersections

Columns:

- FireDisturbCodedOnPlot:
  - False: disturbance NOT coded for any condition at any visit under annual inventory
  - True: disturbance WAS coded but for a pre-1960 fire (SO, does not mean that a post-1960 disturbance was coded)

```{r}

intrsct_dat <- read_excel("../data/FireOverlay2.xlsx") %>%
  arrange(plot_fiadb, FireYear) %>%
  rename(PerimFireYear = FireYear)

```

```{r}

joined_dat <- q2 %>%
  left_join(intrsct_dat, by = c("PLOT_FIADB" = "plot_fiadb"), keep = T) %>%
  select(PLOT_FIADB, INVYR, DSTRBCD1, CONDID, DSTRBYR1, Source, TimeBtwnFires, PerimFireYear) %>%
  group_by(PLOT_FIADB) %>%
  mutate(match = DSTRBYR1 %in% PerimFireYear,
         n = n_distinct(match),
         n_FIAfires = n_distinct(DSTRBYR1),
         nPerimFires = n_distinct(PerimFireYear)) %>%
  ungroup()

rcnt_jnd_dat <- joined_dat %>%
  filter(DSTRBYR1 > 1960) %>%
  group_by(PLOT_FIADB) %>%
  mutate(n_FIAfires = n_distinct(DSTRBYR1)) %>%
  filter(n_FIAfires >= 2) %>%
  mutate(match = DSTRBYR1 %in% PerimFireYear,
         n = n_distinct(match)) %>%
  mutate(DstncBtwnFireYrs = abs(DSTRBYR1 - PerimFireYear)) %>%
  ungroup()

length(unique(joined_dat$PLOT_FIADB))
length(unique(rcnt_jnd_dat$PLOT_FIADB))

```

#### Build a helper function to assign observations to "keep" vs. "check" lists

```{r}

plts2kp <- c()
plts2chk <- c()

assignPlots <- function(df) {
  
  keepPlts <- df %>%
    filter(keep == "keep") %>%
    pull(PLOT_FIADB) %>%
    unique() 
  
  chkPlts <- df %>%
    filter(keep == "check") %>%
    pull(PLOT_FIADB) %>%
    unique() 
  
  plts2kp <<- c(plts2kp, keepPlts)
  plts2chk <<- c(plts2chk, chkPlts)

  
}

```


### Confirmed plots table

```{r}

confirmed <- rcnt_jnd_dat %>%
  filter(match == TRUE, n == 1) %>%
  group_by(PLOT_FIADB) %>%
  mutate(n_FIAfires = n_distinct(DSTRBYR1)) %>%
  ungroup() %>%
  filter(match == TRUE, n == 1, n_FIAfires == 2) %>%
  mutate(keep = "keep")

length(unique(confirmed$PLOT_FIADB))

assignPlots(confirmed)

```

### Partly confirmed plots table

```{r}

prtly_cnfrmd <- rcnt_jnd_dat %>%
  filter(n == 2) %>%
  group_by(PLOT_FIADB) %>%
  mutate(n_FIAfires = n_distinct(DSTRBYR1)) %>%
  ungroup()

length(unique(prtly_cnfrmd$PLOT_FIADB))

# Rule 1, if there are two burn years and the distance between them is less than
# or equal to 5 years, move to the "to check" category

prtly_p1 <- prtly_cnfrmd %>%
  group_by(PLOT_FIADB) %>%
  mutate(n_invyr = n_distinct(INVYR),
         n_brnyr = n_distinct(DSTRBYR1),
         dstnc_brnyrs = abs(max(DSTRBYR1) - min(DSTRBYR1)),
         keep = ifelse(n_brnyr == 2 & dstnc_brnyrs <= 5, "check", NA)) %>%
  ungroup()

assignPlots(prtly_p1)

# Rule 2, if the plot has multiple INVYRs, and the historical fire
# disturbance is confirmed in the second visit, keep the plots

prtly_p2 <- prtly_p1 %>%
  filter(is.na(keep)) %>%
  group_by(PLOT_FIADB, Source) %>%
  mutate(n_src = ifelse(Source == "HIST_DISTURB", n_distinct(Source), NA),
         n_brnyr = n_distinct(DSTRBYR1),
         n_invyr = n_distinct(INVYR)) %>%
  mutate(keep = ifelse(n_invyr == 2 & n_src == 1 & n_brnyr == 1, "keep", NA)) %>%
  ungroup() %>%
  group_by(PLOT_FIADB) %>%
  mutate(keep = ifelse("keep" %in% keep, "keep", NA)) %>%
  ungroup()

assignPlots(prtly_p2)

# Rule 3, if the plot was visited twice, there were two burns, and second burn
# occurred between the two visits, keep the plot
# Note: plot 61673 has a brnyr after its initial inventory, which was due to a
# measure year that occurred after the invyr

prtly_p3 <- prtly_p2 %>%
  filter(is.na(keep)) %>%
  group_by(PLOT_FIADB) %>%
  mutate(n_invyr = n_distinct(INVYR),
         n_brnyr = n_distinct(DSTRBYR1),
         keep = ifelse(n_invyr == 2 & n_brnyr == 2 & max(DSTRBYR1) <= max(INVYR) & max(DSTRBYR1) >= min(INVYR), "keep", NA)) 

assignPlots(prtly_p3)

# Rule 4, clear out the 9999 values

prtly_p4 <- prtly_p3 %>%
  filter(is.na(keep)) %>%
  group_by(PLOT_FIADB) %>%
  mutate(keep = ifelse(9999 %in% DSTRBYR1, "check", NA))

assignPlots(prtly_p4)

# Rule 5, if a plot has a single inventory year associated with it but multiple
# burn years that are more than 5 years apart, keep the plot given that the 
# multiple dates are unlikely to point to the same event.

prtly_p5 <- prtly_p4 %>%
  filter(is.na(keep)) %>%
  group_by(PLOT_FIADB, INVYR) %>%
  mutate(nBrnyrsByVisit = n_distinct(DSTRBYR1),
         dstncBtwnBrnYrs = abs(max(DSTRBYR1) - min(DSTRBYR1))) %>%
  ungroup() %>%
  mutate(keep = ifelse(n_invyr == 1 & n_brnyr == 2, "keep", NA))

assignPlots(prtly_p5)

# Rule 6, check remaining plots with 3 burns

prtly_p6 <- prtly_p5 %>%
  filter(is.na(keep)) %>%
  mutate(keep = ifelse(n_brnyr == 3, "check", NA))

assignPlots(prtly_p6)

# Rule 7, if a plot has three inventory years and the 2nd and third inventory years
# report the same disturbance date that falls between visit 1 and visit 2, keep
# the plots
# e.g., inventories in 2001, 2004, and 2014 with disturbance dates of 1990, 2003 & 2003

prtly_p7 <- prtly_p6 %>%
  filter(is.na(keep)) %>%
  group_by(PLOT_FIADB) %>%
  mutate(keep = ifelse(max(DSTRBYR1) < median(INVYR) & max(DSTRBYR1) > min(INVYR), "keep", NA))
  
assignPlots(prtly_p7)

# Rule 8, If a plot has three inventories and the first and second inventory 
# both point to one burn whereas the third inventory points to a second 
# burn that is in between visit 2 and visit 3, keep the plot (Similar logic
# to rule 7, but other "direction")

prtly_p8 <- prtly_p7 %>%
  filter(is.na(keep)) %>%
  group_by(PLOT_FIADB) %>%
  mutate(keep = ifelse(n_brnyr == 2 & max(DSTRBYR1) > median(INVYR), "keep", NA))
  
assignPlots(prtly_p8)

# Rule 9, remaining plots are a bit haphazard, a fair number with >3 inventories, 
# in looking at them, plot 68034 looks okay - move others to the check category

prtly_p9 <- prtly_p8 %>%
  filter(is.na(keep)) %>%
  mutate(keep = ifelse(PLOT_FIADB == 68034, "keep", "check")) %>%
  arrange(PLOT_FIADB, INVYR)

assignPlots(prtly_p9)


#------

length(plts2chk)
length(unique(plts2chk))

length(plts2kp)
length(unique(plts2kp))

```


### Unconfirmed plots table

Handling the unconfirmed data

```{r}

uncnfrmd <- rcnt_jnd_dat %>%
  filter(match == FALSE, n == 1)

length(unique(uncnfrmd$PLOT_FIADB))


uncnfrmd2cln <- uncnfrmd %>%
  group_by(PLOT_FIADB) %>%
  mutate(n_invyr = n_distinct(INVYR),
         n_brnyr = n_distinct(DSTRBYR1),
         n_src = n_distinct(Source),
         dstnc_brnyrs = abs(max(DSTRBYR1) - min(DSTRBYR1))) %>%
  ungroup()


# Rule 1, If 9999 (missing data) is in the fire disturbance year column,
# assign the plot number to the "check" list

uncnfrmd_p1 <- uncnfrmd2cln %>%
  group_by(PLOT_FIADB) %>%
  mutate(keep = ifelse(9999 %in% DSTRBYR1, "check", NA)) %>%
  ungroup()
  
assignPlots(uncnfrmd_p1)

# Rule 2, if only one INVYR and multiple fire disturbances are recorded, keep
# under the assumption that those measuring had the correct information and
# would not create errors of commission.

uncnfrmd_p2 <- uncnfrmd_p1 %>%
  filter(is.na(keep)) %>%
   mutate(keep = ifelse(n_invyr == 1 & n_brnyr == 2, "keep", NA))

assignPlots(uncnfrmd_p2)

# Rule 3, if the distance between the fire disturbances is less than 5 years,
# place the plot in the "to check" list.

uncnfrmd_p3 <- uncnfrmd_p2 %>%
  filter(is.na(keep)) %>%
  mutate(keep = ifelse(dstnc_brnyrs <= 5, "check", NA))

assignPlots(uncnfrmd_p3)

# Rule 4, if the plot has multiple INVYRs, and the historical fire
# disturbance is confirmed in the second visit, keep the plots

uncnfrmd_p4 <- uncnfrmd_p3 %>%
  filter(is.na(keep)) %>%
  group_by(PLOT_FIADB, Source) %>%
  mutate(n_src = ifelse(Source == "HIST_DISTURB", n_distinct(Source), NA),
         n_brnyr = n_distinct(DSTRBYR1),
         n_invyr = n_distinct(INVYR)) %>%
  mutate(keep = ifelse(n_invyr == 2 & n_src == 1 & n_brnyr == 1, "keep", NA)) %>%
  ungroup() %>%
  group_by(PLOT_FIADB) %>%
  mutate(keep = ifelse("keep" %in% keep, "keep", NA)) %>%
  ungroup()

assignPlots(uncnfrmd_p4)

# Rule 5, If a plot was visted twice, and the second visit records only a burn
# between the first and second visit (but not historical burn), keep the plot
# under the assumption that the more recent fire took precedence over the
# (valid) historical fire

uncnfrmd_p5 <- uncnfrmd_p4 %>%
  filter(is.na(keep)) %>%
  group_by(PLOT_FIADB) %>%
  mutate(n_invyr = n_distinct(INVYR),
         n_brnyr = n_distinct(DSTRBYR1),
         keep = ifelse(n_invyr == 2 & max(DSTRBYR1) <= max(INVYR) & max(DSTRBYR1) >= min(INVYR), "keep", NA)) %>%
  ungroup()

assignPlots(uncnfrmd_p5)

# Rule 6, for plots with 2 visits & 2 burn years, but both burn years are 
# before the first inventory, move the plots to the "check" dataframe
# e.g., visited in 2002 and 2006 with 2002 reporting brn yr of 1970 and 
# 2006 reporting brn yr of 1960

uncnfrmd_p6 <- uncnfrmd_p5 %>%
  filter(is.na(keep)) %>%
  group_by(PLOT_FIADB) %>%
  mutate(keep = ifelse(n_invyr == 2 & n_brnyr == 2 & max(DSTRBYR1) < min(INVYR) & max(DSTRBYR1) != min(DSTRBYR1), "check", NA))

assignPlots(uncnfrmd_p6)

# Rule 7, remaining plots are a bit haphazard, manually sort through them saving
# four plots and assigning the rest to the "check" category

uncnfrmd_p7 <- uncnfrmd_p6 %>%
  filter(is.na(keep)) %>%
  group_by(PLOT_FIADB) %>%
  mutate(keep = ifelse(PLOT_FIADB %in% c(68381, 68385, 55140, 77768), "keep", "check")) %>%
  arrange(PLOT_FIADB, INVYR)

assignPlots(uncnfrmd_p7)

# View final results

length(plts2kp)
length(unique(plts2kp))

length(plts2chk)
length(unique(plts2chk))

```

```{r}

write_csv(plts2chk)
write_csv(plts2kp)

```